Supplementary material: The increasing cost of happiness

### 2021-03-12 src/supplementary0.Rmd
Submitted to PNAS

---
title: "Supplementary material: The increasing cost of happiness"
author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,5^"
bibliography: ../src/references.bib
csl: ../src/elsevier-harvard.csl
output:  
  html_document:
    reference_docx: ../src/academic_style.docx
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

@@

1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
2. ARC Centre of Excellence for Children and Families over the Life Course, University of Queensland, QLD, Australia
3. School of Economics, University of Technology, NSW, Australia
4. Institute of Labor Economics (IZA), Bonn, Germany
5. Brain and Mind Centre, University of Sydney, NSW, Australia

@@

##### Figure S1. Linear model evidence (WAIC~linear~ — WAIC~piecewise~)
```{r figure_s1, dpi=300}
here <- "~/Dropbox/HILDA/The-increasing-cost-of-happiness/"

lm_fits = read_rds(paste0(here, "results/lm_gh9_hhoecd_0cov.RDS"))

nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight.RDS"))

@@

nl_fits_losat = read_rds(paste0(here, "results/losat_hhoecd_0cov.RDS"))

lm_fits_losat = read_rds(paste0(here, "results/lm_losat_hhoecd_0cov.RDS"))

@@

##### Figure S2. Conditional model parameters  
```{r figure_s2, fig.height=8, dpi=300}
here <- "~/Dropbox/HILDA/The-increasing-cost-of-happiness/"

nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight.RDS"))

nl_fits_5cov = read_rds(paste0(here, "results/gh9_hhoecd_5cov_tight.RDS"))

income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")

posteriors_5c <- map_dfr(nl_fits_5cov, as.data.frame, .id = "fit") %>%
  left_join(bind_rows(income_params, .id = "fit")) %>%
  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
  as_tibble()

@@

posteriors_0c <- map_dfr(nl_fits, as.data.frame, .id = "fit") %>%
  left_join(bind_rows(income_params, .id = "fit")) %>%
  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
  as_tibble()

@@

##### Figure S3. Log-linear model evidence (WAIC~log-linear~ — WAIC~piecewise~)
```{r, figure_s3, fig.height=3.5, dpi=300}
here <- "~/Dropbox/HILDA/The-increasing-cost-of-happiness/"

nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight_x0.RDS"))

log_fits = read_rds(paste0(here, "results/log_gh9_hhoecd_0cov.RDS"))

# get waic results
map2_dfr(.x = nl_fits, 
         .y = log_fits, 
         .f = ~{
           loo_compare(.x, .y, criterion = "waic") %>%
             as.data.frame() %>%
             rownames_to_column("model") %>%
             as_tibble()
         },
         .id = "year") %>%
  filter(elpd_diff != 0) %>%
  mutate(
    year = extract_numeric(year),
    model = if_else(model == ".x", "nl", "log-linear"),
    waic_diff = 2 * elpd_diff, # lm disadvantage
    se = 2 * se_diff
  ) %>%
  ggplot(aes(x = waic_diff, y = year)) +
  geom_vline(aes(xintercept = 0), color = "grey90") +
  geom_pointrange(aes(xmin = waic_diff - se*1.96, xmax = waic_diff + se*1.96),
                 color = "#ffcc80") +
  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
  ft_theme() +
  labs(x = "deviance", y = "")
```
```
Figure S3 legend: The joint distribution between wellbeing and income for a single year, imposed by the priors.
```

@@

##### Figure S4. Prior predictive check  
```{r figure_s4, dpi=300}
load(paste0(here, "results/prior_predictions.Rdata"))

@@

load(paste0(here, "results/residual_fits.Rdata"))

@@

```{r table_s2, ft.align="left"}
param_key = c(
  b0_Intercept = "Intercept",
  b1_Intercept = "Pre-slope",
  b2_Intercept = "Post-slope",
  omega_Intercept = "Change-point"
)

map_dfr(nl_fits, 
        .f = ~{
          fit_summary <- summary(.x)

@@

          },
        .id = "Year") %>%
  mutate(
    Year = str_remove(Year, "fit"),
    Parameter = recode_factor(Parameter, !!!param_key)
    ) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Year %in% seq(2001, 2019, 2), 
     bg = "grey90", 
     part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S2. Piecewise parameter estimates and fit statistics for happiness") %>%
  autofit()
```

@@

          },
        .id = "Year") %>%
  mutate(
    Year = str_remove(Year, "fit"),
    Parameter = recode_factor(Parameter, !!!param_key)
    ) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Year %in% seq(2001, 2019, 2), 
     bg = "grey90", 
     part = "body") %>%
  colformat_num(j = 3:6, digits= 3) %>%
  colformat_num(j = 7, digits= 1) %>%
  set_caption("Table S3. Linear parameter estimates and fit statistics for life satisfaction") %>%
  autofit()
```

@@

##### Figure S6. Happiness (SD units) at $50K/yr and $75K/yr from 2001-2019  
```{r figure_s6, fig.width=7, fig.height=5, dpi=300}
here <- "~/Dropbox/HILDA/The-increasing-cost-of-happiness/"

nl_fits = read_rds(paste0(here, "results/gh9_hhoecd_0cov_tight.RDS"))

@@

income_params <- read_rds("../results/gh9_hhoecd_summary.RDS")

bind_rows(income_params, .id = "fit") %>%
  mutate(low = (40000 - mean)/sd,
         high = (75000 - mean)/sd) %>%
  column_to_rownames(var = "fit") %>%
  as.matrix() -> fixed_income_hash

@@

model_summaries %>%
  select(model, year, r.squared) %>%
  spread(model, r.squared) %>%
  flextable() %>%
  colformat_num(j = 2:7, digits = 3) %>%
  set_caption("Table S4. In-sample fit statistics (R-squared)") %>%
  set_header_labels(
    lin = "Linear", 
    linc = "Linear\n(conditional)",
    log = "Log-linear",
    logc = "Log-linear\n(conditional)",
    pw = "Piecewise",
    pwc = "Piecewise\n(conditional)") %>%
  valign(i = 1, valign = "top", part = "header") %>%
  width(j = 2:7, width = 1)
```

@@

Three things need to be observed when interpreting the R^2^ values: 1) While the R^2^ between income and happiness is low (~2% and ~14% in the unconditional and conditional models, respectively), the aggregate effect over the wider population (e.g., 25M in Australia) will be much larger; 2) Happiness (and to a lesser extent, income) is an imperfect and noisy measure so any effect of income on happiness is likely to be larger than that measured here; 3) Finally, happiness is measured on a subjective and arbitrary Likert scale so a small change in happiness at different points in the scale may represent a large change in some other functional/real-world outcome (e.g., risk of suicide). This is all to say that quantifying the real-world impact of small changes in variation reported here is difficult and speculative without further investigation.

>>>>

---
title: "Supplementary material: The increasing cost of happiness"
author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,2^"
bibliography: ../src/references.bib
csl: ../src/elsevier-harvard.csl
output:  
  html_document:
    reference_docx: ../src/academic_style.docx
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

@@

1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
2. ARC Centre of Excellence for Children and Families over the Life Course, University of Queensland, QLD, Australia
3. School of Economics, University of Technology, NSW, Australia
4. Institute of Labor Economics (IZA), Bonn, Germany

@@

##### Figure S1. Linear model evidence (WAIC~linear~ — WAIC~piecewise~)
```{r figure_s1, dpi=300}
lm_fits = read_rds("../results/lm_gh9_hhoecd_0cov.RDS")

nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")

@@

nl_fits_losat = read_rds("../results/losat_hhoecd_0cov.RDS")

lm_fits_losat = read_rds("../results/lm_losat_hhoecd_0cov.RDS")

@@

##### Figure S2. Conditional model parameters  
```{r figure_s2, fig.width = 6, fig.height = 7.5, dpi=300}
nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")

nl_fits_5cov = read_rds("../results/gh9_hhoecd_5cov_tight.RDS")

income_params <- read_rds("../results/gh9_hhoecd_summary.RDS") %>%
  bind_rows(.id = "fit")

posteriors_5c <- map_dfr(nl_fits_5cov, as.data.frame, .id = "fit") %>%
  left_join(income_params) %>%
  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
  as_tibble()

@@

posteriors_0c <- map_dfr(nl_fits, as.data.frame, .id = "fit") %>%
  left_join(income_params) %>%
  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
  as_tibble()

@@

##### Figure S3. Log-linear model evidence (WAIC~log-linear~ — WAIC~piecewise~)
```{r, figure_s3, fig.height=3.5, dpi=300}
nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight_x0.RDS")

log_fits = read_rds("../results/log_gh9_hhoecd_0cov.RDS")

# get waic results
map2_dfr(.x = nl_fits, 
         .y = log_fits, 
         .f = ~{
           loo_compare(.x, .y, criterion = "waic") %>%
             as.data.frame() %>%
             rownames_to_column("model") %>%
             as_tibble()
         },
         .id = "year") %>%
  filter(elpd_diff != 0) %>%
  mutate(
    year = extract_numeric(year),
    model = if_else(model == ".x", "nl", "log-linear"),
    waic_diff = 2 * elpd_diff, # lm disadvantage
    se = 2 * se_diff
  ) %>%
  ggplot(aes(x = waic_diff, y = year)) +
  geom_vline(aes(xintercept = 0), color = "grey90") +
  geom_pointrange(aes(xmin = waic_diff - se*1.96, xmax = waic_diff + se*1.96),
                 color = "#ffcc80") +
  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
  ft_theme() +
  labs(x = "deviance", y = "")
```
```
Figure S3 legend: Differences in WAIC scores (deviance units) for a log-linear fit over a piecewise fit for happiness.
```

@@

##### Figure S4. Prior predictive check  
```{r figure_s4, dpi=300}
load("../results/prior_predictions.Rdata")

@@

load("../results/residual_fits.Rdata")

@@

```{r table_s2, ft.align="left"}
nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")

income_params %>%
  mutate(fit = str_remove(fit, "fit")) %>%
  column_to_rownames("fit") -> income

param_key = c(
  b0_Intercept = "Intercept",
  b1_Intercept = "Pre-slope",
  b2_Intercept = "Post-slope",
  omega_Intercept = "Change-point ($000s)"
)

map_dfr(nl_fits, 
        .f = ~{
          fit_summary <- summary(.x, prob = 0.95)

@@

          },
        .id = "Year") %>%
  mutate(
    Year = str_remove(Year, "fit"),
    Parameter = recode_factor(Parameter, !!!param_key),
    Estimate = if_else(str_detect(Parameter, "Change"),
                       (Estimate * income[Year, "sd"] + income[Year, "mean"])/1000,
                       Estimate),
    Est.Error = if_else(str_detect(Parameter, "Change"),
                        (Est.Error * income[Year, "sd"] + income[Year, "mean"])/1000,
                        Est.Error),
    `l-95% CI` = if_else(str_detect(Parameter, "Change"),
                        (`l-95% CI` * income[Year, "sd"] + income[Year, "mean"])/1000,
                        `l-95% CI`),
    `u-95% CI` = if_else(str_detect(Parameter, "Change"),
                         (`u-95% CI` * income[Year, "sd"] + income[Year, "mean"])/1000,
                         `u-95% CI`)
    ) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Year %in% seq(2001, 2019, 2), 
     bg = "grey90", 
     part = "body") %>%

  colformat_double(j = 3:6, digits= 3) %>%
  colformat_double(j = 7, digits= 1) %>%
  set_caption("Table S2. Piecewise parameter estimates and fit statistics for happiness") %>%
  fontsize(size = 9.5, part = "all") %>%
  width(j = 1, width = .5) %>%
  width(j = 2, width = 1) %>%
  width(j = c(3:7), width = .7) %>%
  width(j = c(8, 9), width = .8)
```

@@

          },
        .id = "Year") %>%
  mutate(
    Year = str_remove(Year, "fit"),
    Parameter = recode_factor(Parameter, !!!param_key)
    ) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Year %in% seq(2001, 2019, 2), 
     bg = "grey90", 
     part = "body") %>%
  colformat_double(j = 3:6, digits= 3) %>%
  colformat_double(j = 7, digits= 1) %>%
  set_caption("Table S3. Linear parameter estimates and fit statistics for life satisfaction") %>%
  fontsize(size = 9.5, part = "all") %>%
  autofit()
```

@@

##### Figure S6. Happiness (SD units) at $50K/yr and $75K/yr from 2001-2019  
```{r figure_s6, fig.width=7, fig.height=5, dpi=300}
nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")

@@

fixed_income_hash <- read_rds("../results/gh9_hhoecd_summary.RDS") %>%
  bind_rows(.id = "fit") %>%
  mutate(low = (40000 - mean)/sd,
         high = (75000 - mean)/sd) %>%
  column_to_rownames(var = "fit") %>%
  as.matrix()

@@

model_summaries %>%
  select(model, Year = year, r.squared) %>%
  spread(model, r.squared) %>%
  flextable() %>%
  colformat_num(j = 1, big.mark = "") %>%
  colformat_double(j = 2:7, digits = 3) %>%
  set_caption("Table S4. In-sample fit statistics (R-squared)") %>%
  set_header_labels(
    lin = "Linear", 
    linc = "Linear\n(conditional)",
    log = "Log-linear",
    logc = "Log-linear\n(conditional)",
    pw = "Piecewise",
    pwc = "Piecewise\n(conditional)") %>%
  valign(i = 1, valign = "top", part = "header") %>%
  fontsize(size = 9.5, part = "all") %>%
  width(j = 2:7, width = 1) %>%
  width(j = 1, width = .5)
```

@@

Three things need to be observed when interpreting the R^2^ values: 1) While the R^2^ between income and happiness is low (~2% and ~14% in the unconditional and conditional models, respectively), the effect over the wider population will aggregate to be much larger; 2) Happiness (and to a lesser extent, income) is an imperfect and noisy measure so any effect of income on happiness is likely to be larger than that measured here; 3) Finally, happiness is measured on a subjective and arbitrary Likert scale so a small change in happiness at different points in the scale may represent a large change in some other functional/real-world outcome (e.g., risk of suicide). This is all to say that quantifying the real-world impact of small changes in variation reported here is difficult and speculative without further investigation.

### 2021-03-12 src/supplementary0_word.Rmd
Submitted to PNAS

---
title: "Supplementary material: The increasing cost of happiness"
author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,5^"
bibliography: ../src/references.bib
csl: ../src/elsevier-harvard.csl
output:  
  html_document:
    reference_docx: ../src/academic_style.docx
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

@@

1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
2. ARC Centre of Excellence for Children and Families over the Life Course, University of Queensland, QLD, Australia
3. School of Economics, University of Technology, NSW, Australia
4. Institute of Labor Economics (IZA), Bonn, Germany
5. Brain and Mind Centre, University of Sydney, NSW, Australia

>>>>

---
title: "Supplementary material: The increasing cost of happiness"
author: "R.W. Morris^1,2^, N. Kettlewell^2,3,4^  & N.Glozier^1,2^"
bibliography: ../src/references.bib
csl: ../src/elsevier-harvard.csl
output:  
  word_document:
    reference_docx: ../src/Rmarkdown-for-Word-styles.docx
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

@@

1. Central Clinical School, Faculty of Medicine and Health, University of Sydney, NSW, Australia
2. ARC Centre of Excellence for Children and Families over the Life Course
3. School of Economics, University of Technology, NSW, Australia
4. Institute of Labor Economics (IZA), Bonn, Germany
